---
title: "Chapter 02. The Insurance Market and Our Case Studies"
author: "Zach Eisenstein"
date: "`r format(Sys.Date())`"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

To Do:

- remove dependencies to internal formatting (AonECM, AonAltair)


```{r, warning = F, message = F}
library(tidyverse)
library(flextable)
theme_set(theme_light())
```

## 2.4.1 The Simple Discrete Example

We start with a simple discrete distribution on a two unit portfolio Let's warm up a bit by inputting and getting some basic statistics.


```{r}
x1 <- tibble(x1_gross = c(0, 8, 10),
             x1_net = c(0, 8, 10),
             px1 = c(.5, .25, .25))
```


X2 has reinsurance applied that caps losses at 20.

```{r}
x2 <- tibble(x2_gross = c(0, 1, 90),
             x2_net = c(0, 1, 20),
             px2 = c(.5, .25, .25))
```

We'll bring the distributions together using the `crossing` function and multiply probabilities (due to independence assumption).

```{r}
x_total <- crossing(x1, x2) %>%
  mutate(total_gross = x1_gross + x2_gross,
         total_net = x1_net + x2_net,
         p = px1 * px2) %>%
  select(-px1, -px2)

x_total
```

Let's plot the distributions

```{r}
x_total %>%
  pivot_longer(-p,names_to = c("Line", "GN"), names_sep = "_") %>%
  ggplot(aes(value, p)) +
  geom_col() +
  facet_grid(GN ~ Line)
```

...and calculate summary statistics.

We'll use a helper function for this.

```{r}
discrete_sum <- function(x, p){
  
  .moment <- function(x, p, k = 1) {
    
    m = sum(x * p)
    sum((x - m) ^ k * p)
  }
  
  tibble(Mean = sum(x * p),
         CV = (.moment(x, p, 2) ^ .5) / Mean,
         Skewness = .moment(x, p, 3) / .moment(x, p , 2) ^ (3/2),
         Kurtosis = .moment(x, p, 4) / .moment(x, p , 2) ^ 2 - 3
  )
  
  
}
```

These figures align with Table 2.3 (p19)

```{r}
x_total %>%
  pivot_longer(-p,names_to = c("Line", "GN"), names_sep = "_") %>%
  group_by(GN, Line) %>%
  nest() %>%
  summarise(s = map(data, ~ discrete_sum(.x$value, .x$p))) %>%
  ungroup() %>%
  unnest(s) %>%
  pivot_longer(Mean:Kurtosis, names_to = "Statistic") %>%
  mutate(Line = factor(Line, levels = c('x1', 'x2', 'total'))) %>%
  arrange(GN, Line) %>%
  pivot_wider(names_from = c(GN, Line)) %>%
  mutate_if(is.numeric, ~ round(., 3)) %>%
  flextable::flextable() %>%
  flextable::set_header_labels(Statistics = "Statistic",
                               gross_total = "Total",
                               gross_x1 = "X1",
                               gross_x2 = "X2",
                               net_total = "Total",
                               net_x1 = "X1",
                               net_x2 = "X2") %>%
  flextable::add_header_row(
    values = c("","Gross", "Net"),
    colwidths = c(1, 3, 3)) %>%
  AonECM::ft_theme_aon() %>%
  AonECM::add_title_header("Table 2.3")
  # flextable::align(align = "left", part = "header") %>%
  # add_header_lines("Table 2.3") %>%
  # fontsize(i = 1, size = 14, part = "header") %>%
  # flextable::border(i = 1, border.top = officer::fp_border(color = "transparent"))
```

If we wanted to sample from the discrete distribution we could use `replicate` or `slice_sample`

```{r}

x_total %>% 
  slice_sample(n = 1e3, weight_by = p, replace = T) %>%
  select(-p)
```

## 2.4.2 Tame Case Study

> In the Tame Case Study, Ins Co. write two predictable units with no catastrophe exposure. We include it to demonstrate an idealized risk-pool: it represents the best case--from Ins Co.'s perspective. It could proxy a portfolio of personal and commercial auto liability. 

For this case study we have two units, A & B that are each [gamma](https://en.wikipedia.org/wiki/Gamma_distribution) distributed with mean 50 and CV 10% and 15% respectively. 
The first step is to express these assumptions as parameters to the gamma distribution

With shape parameter $\alpha$ and scale $\sigma$, the mean and variance are $E[X] = \alpha  \sigma$ and $Var[X] = \alpha\sigma^{2}$

This implies that $\alpha = \frac{1}{CV[X]^{2}}$ and $\sigma = E[X]CV[X]^{2}$ or $\sigma = \frac{E[X]}{\alpha}$

When using the `rate` instead of `scale`, we'd simply use $\frac{1}{\sigma}$

```{r}

tame <- tribble(
  ~ Unit, ~ Distribution, ~ Mean, ~ CV,
  "A", "Gamma", 50, .1,
  "B", "Gamma", 50, .15
) %>%
  mutate(alpha = 1 / CV^2) %>%
  mutate(rate = alpha / Mean)

tame %>%
  mutate(across(c(alpha, rate), ~ round(., 3))) %>%
  knitr::kable(caption = "Tame Case Study Parameters")


```
It's noted that aggregate reinsurance is applied to B with an attachment probability of 0.2 and exhaust of .01. 
Let's calculate the precise limit and attachments

```{r}
tame_ri <- tibble(unit = "B",
                  attach = qgamma(1 - .2, tame$alpha[2], tame$rate[2]),
                  exhaust = qgamma(1 - .01, tame$alpha[2], tame$rate[2]),
                  limit = exhaust - attach)
tame_ri
```




Let's run a simulation

```{r}
sims <- tame %>%
  crossing(trial = 1:1e6) %>%
  mutate(value = rgamma(n(), alpha, rate)) %>%
  select(trial, Unit, value) %>%
  pivot_wider(names_from = Unit) %>%
  mutate(Total = A + B)

# add reinsurance on B
sims <- sims %>%
  rename_with(~ paste0(., "_Gross"), .cols = -trial) %>%
  mutate(A_Net = A_Gross) %>%
  mutate(B_Net = B_Gross - pmin(tame_ri$limit, pmax(B_Gross - tame_ri$attach, 0))) %>%
  mutate(Total_Net = A_Net + B_Net)
  
head(sims)
```

The results should be comparable to Table 2.5 (p20), but will differ due to simulation error

```{r}
sims %>%
  select(-trial) %>%
  summarise_all(list(Mean = mean,
            CV = ~ sd(.) / mean(.),
            Skewness = moments::skewness,
            Kurtosis = ~ moments::kurtosis(.) - 3)) %>%
  pivot_longer(everything(), names_to = c("Line", "GN", "Statistic"), names_sep = "_") %>%
  pivot_wider(names_from = c(GN, Line)) %>%
  mutate(across(is.numeric, ~ round(., 3)))
```


We could use Fast Fourier Transform as Mildenhall does. 
> The FFT approximation is essentially exact and it is used for all graphics and illustrations. 


```{r}
buckets <- 2^18
pml <- 200 # maximum of discretized distribution
step_size <- pml / buckets

tame_fft <- tibble(start = 0:(buckets-1) * step_size,
       mid = start + step_size / 2) %>%
  mutate(A = pgamma(mid, tame$alpha[1], tame$rate[1])) %>%
  mutate(B = pgamma(mid, tame$alpha[2], tame$rate[2])) %>%
  
  # differences of Fx
  mutate(across(c(A, B), ~if_else(row_number() == 1, ., . - lag(.)))) %>%
  
  # Aggregate approximation from FFT
  mutate(Total = Re(fft(fft(A) * fft(B), inverse = T)) / buckets)
  
tame_fft

tame_fft %>%
  select(-start) %>%
  pivot_longer(-mid) %>%
  group_by(name) %>%
  summarise(discrete_sum(mid, value)) %>%
  pivot_longer(-name, names_to = "Statistic") %>%
  pivot_wider() %>%
  mutate(across(is.numeric, ~ round(., 3)))

```

Attempt at making CDF of B Net
```{r}
tame_fft

```


Let's take a look at the densities

```{r}

tame %>%
  select(Unit, alpha, rate) %>%
  crossing(x = seq(0, 200, length.out = 100)) %>%
  mutate(d = dgamma(x, alpha, rate, log = F)) %>%
  ggplot(aes(x, d, color = Unit)) +
  geom_line()

ggplot(data = data.frame(x = c(0, 200)), aes(x)) +
  stat_function(fun = dgamma, n = 101, args = list(shape = tame$Mean[1], rate = tame$rate[1])) + ylab("") +
  scale_y_continuous(breaks = NULL)

```

